Sulfur Dioxide on Water's Surface

### 2011-06-07 Conclusions.tex
Finished first pass of fixing up draft 4.

@@

Our simulations recreated the experimental setup of a recent study by our group,\cite{Ota2011} by introducing a concentrated gas phase of \suldiox~to a water surface. It was found that gaseous \suldiox~quickly adsorbs to the water surface, and continues to bind until a complete surface coverage is reached. Surface waters reorient in the presence of adsorbed \suldiox. The waters at the interface of a neat-water surface tend to lay flatter to the surface than when a saturating layer of \suldiox~is present. The waters interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation.

The surface behavior of the \suldiox~molecules was also determined. Vibrational sum-frequency surface spectroscopy on the \suldiox~vibrational spectra is not possible because of the nature of the experimental geometry. At aqueous interfaces, the spectra of the surface \suldiox~vibrational modes are still unreported. However, through these simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption. Our equilibrium simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity. This was previously reported by Baer et al.\cite{Baer2010} At a high \suldiox~concentration, unreacted \suldiox~molecules are found further out of the water phase as a bound layer of \suldiox~crowds the surface and binds all the surface waters. The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations. Those \suldiox~molecules at or below the surface strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase. The \suldiox~slightly above the water surface loses the orientational preference within 10 $\AA$ and those further from the water are more isotropically oriented. Figure \ref{fig:so2-surface-cartoon} depicts what the neat-water and saturated surfaces might look like for both \suldiox~and \wat~orientations and locations.

>>>>

% intro-type paragraph here
The aqueous chemistry taking place in the Earth's atmosphere is goverened by the physics of the water surfaces found there. Many recent experiments have revealed specific pathways for reaction of \suldiox, but the molecular-level nature of those reactions are still poorly understood. Through computer simulations we are able to probe the microscopic environments in which those reactions occur, and better understand the conditions in which \suldiox~reacts.

@@

Our simulations recreated the experimental setup of a recent study by our group,\cite{Ota2011} by introducing a concentrated gas phase of \suldiox~to a water surface. Through simulation it was found that gaseous \suldiox~quickly adsorbs to the water surface, and continues to bind until a complete surface coverage is reached. Surface waters reorient in the presence of adsorbed \suldiox. The waters at the interface of a neat-water surface tend to lay flatter to the surface than when a saturating layer of \suldiox~is present. The waters interacting with the layer of adsorbed \suldiox~orient more perpendicularly to the interface, and further expose their ``free-OH'' uncoupled bonds for interactions with \suldiox, and hydrate complex formation.

The surface behavior of the \suldiox~molecules was also determined, information that is difficult to obtain with vibrational sum-frequency spectroscopy. Through these simulations we are able to characterize the orientational behavior of \suldiox~during and after adsorption. Our equilibrium simulations show that a single \suldiox~molecule, representing a low concentration, has a high surface affinity. At a high \suldiox~concentration, unreacted \suldiox~molecules are found further out of the water phase as a bound layer of \suldiox~crowds the surface and binds all the surface waters. The orientation of \suldiox~on the water surface was found to be similar for both low and high concentrations. Those \suldiox~molecules at or below the surface strongly orient with the sulfur atom pointed in towards the water bulk, and the oxygen atoms out towards the gas phase. The \suldiox~slightly above the water surface loses the orientational preference within 10 $\AA$ and those further from the water are more isotropically oriented. Figure \ref{fig:so2-surface-cartoon} depicts what the neat-water and saturated surfaces might look like for both \suldiox~and \wat~orientations and locations.

### 2011-06-07 EquilibratedOrientation.tex
Finished first pass of fixing up draft 4.

@@

The orientation depth-profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations. In both systems the strongest orientational preference is found at the slab surfaces where the water is furthest towards the gas phase. The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right columns. The surface is located at 0\angs. The angle distributions from both simulated slab surfaces were averaged for all the orientation analyses.

The bisector tilt, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~of the surface, and then the distribution becomes isotropic further into the water bulk. As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a peak in the distributions around $\cos(\phi)=1$ at the water surfaces. This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface. Thus, both the neat-water and saturated systems have waters lying mostly flat to the plane of the interface at a distance of 0\angs.

Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the water orientation of the waters furthest into the gas phase. The waters located above 0\angs~are further into the gas phase, and interact with the layer of adsorbed \suldiox~molecules. The \suldiox~layer does not provide the same amount of strong hydrogen bonding and solvation that the waters experience further in the aqueous phase. The resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk. The effect is more pronounced further from the water phase, and above 5\angs~the $\cos(\theta)$ distribution is completely centered around $\cos(\theta)=+1$ (see figure \ref{fig:water-angles}).

@@

The change in $\cos(\theta)$ is not as rapid at the neat-water surface as in the saturated system. Furthermore, the noise in the distribution above 5\angs~in the neat-water system is a result of fewer waters venturing beyond those extents and thus less data far from the surface. This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface and interact with the adsorbed gas molecules. The interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicular to the interface. Our experimental SFG studies also predicted this reorienting behavior due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/h2o-angles/h2oangles.png}
		\caption{Molecular orientation histograms of \wat~throughout the surface equilibrated systems. The top surface is located at a distance of 0 with negative positions in the bulk of the slab. The bottom slab surface is approx. 30\angs~below the top surface. Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row). The distributions are normalized to account for the changing number of water molecules at different positions in the system.}
		\label{fig:water-orientation}
	\end{center}
\end{figure}

@@

Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-water and saturated systems. Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similarly to the water orientation distributions plots in figure \ref{fig:water-orientation}). The \suldiox~orientation data set for the neat-water system is much smaller as only a single \suldiox~molecule was simulated in the bulk. The resulting distribution plots are thus noisier than the corresponding saturated plots, but effective comparisons can still be drawn.

Within 5\angs~of the water surface (the approximate distance from the surface that \suldiox~begins interacting with the outer-most waters) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$. The peak of the distribution occurs at $\cos(\theta)=1$. The \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase. This same distribution occurs in the saturated system for positions below 5\angs. Beyond 5\angs~above the surface both distributions become isotropic. Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk. Moving further away from the water surface, and interacting less with \wat~molcules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 5\angs. %Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.

@@

>>>>

The 2D orientational depth profile plots show areas of low intensity in dark blue, and highest intensity in dark red. Regions of the plots where the intensity (coloration) is equally distributed from top to bottom along the entire orientational range are considered isotropic. Likewise, areas of the plot with high intensity over a small orientational range are considered to exhibit an orientational preference at the given depth.  The histograms are arranged with the plots of $\cos(\theta)$ on the left and $\cos(\phi)$ on the right columns. The surface is located at 0\angs. The angle distributions from both simulated slab surfaces were averaged for all the orientation analyses.

@@

The orientation depth-profiles for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water (top) and saturated (bottom) systems during the equilibrium MD simulations. In both systems the strongest orientational preference is found at the slab surfaces (distances greater than 0\angs) where the water is furthest towards the gas phase. The narrow "penninsula" of high intensity that extends in the $\cos(\theta)$ plots to the right of the -5\angs~location shows the overall preference of water to orient at the surface. The $\cos(\phi)$ plots are similar to each other with a narrow region of reorientation, but the effect of the surface is greater in the neat-water system as evidenced by the sharper transition in intensity from blue to red, compared to the saturated system that has a less pronounced intensity change.

The bisector tilt of the water molecules, $\cos(\theta)$, concentrates around $\cos(\theta)=0$ within the first few\angs~above the surface, and then the distribution becomes isotropic further into the water bulk of both systems. As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a defined intensity concentration (dark red) in the distributions around $\cos(\phi)=1$ at the water surfaces. This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface. Thus, both the neat-water and saturated systems have waters lying mostly flat to the plane of the interface at a distance of 0\angs.

Although the plots show overall similarities for both the neat-water and saturated systems, the presence of a layer of adsorbed \suldiox~molecules alters the water orientation of the waters furthest into the gas phase. The waters located above 0\angs~are further into the gas phase, and interact with the layer of adsorbed \suldiox~molecules. The resulting orientation of waters above 0\angs, shown in the saturated plot of $\cos(\theta)$ of figure \ref{fig:water-orientation}, is with a bisector pointing further into the adsorbed \suldiox~gas layer, and both hydrogens pointing outward from the aqueous bulk. The effect is more pronounced further from the water phase, and above 5\angs~the $\cos(\theta)$ distribution is completely centered around $\cos(\theta)=+1$ (see figure \ref{fig:water-angles}).

@@

The noise in the neat-water plots of figure \ref{fig:water-orientation} above 5\angs~(manifested as disconnected points of intensity throughout the range of orientations) is a result of fewer waters venturing beyond those extents and thus less data far from the surface. This is one indication that waters near a layer of adsorbed \suldiox~venture further above the interface and interact with the adsorbed gas molecules. The interactions with neighboring \suldiox~molecules allow the waters above the surface to orient more perpendicularly to the interface. Our experimental SFG studies also predicted this reorienting behavior due to the \suldiox~interactions with the topmost surface waters.\cite{Ota2011}

@@

\begin{figure}[h!]
	\begin{center}
		\includegraphics[scale=1.0]{images/h2o-angles/h2oangles.png}
		\caption{Molecular orientation histograms of \wat~throughout the surface equilibrated systems. The top surface is located at a distance of 0\angs~with negative distance values located in the bulk of the slab. Shown are the angle distributions for $\theta$ (left column) and $\phi$ (right column) in both the neat-\wat~system (top row) and the saturated system (bottom row). The distributions are normalized to account for the changing number of water molecules at different positions in the system. Regions of high intensity are dark red, and low intensity are dark blue. The scattered points of coloration to the far right of each plot indicates that few waters were located at those positions, and thus few data points were collected.}
		\label{fig:water-orientation}
	\end{center}
\end{figure}

@@

Orientation distributions of the adsorbed \suldiox~molecules were created during the equilibrium simulations for both the neat-water and saturated systems. Figure \ref{fig:so2-orientation} shows the distributions of $\cos(\theta)$ and $\cos(\phi)$ (arranged similarly to the water orientation distributions plots in figure \ref{fig:water-orientation}). The \suldiox~orientation data set for the neat-water system is much smaller as only a single \suldiox~molecule was simulated in the bulk. The resulting distribution plots are thus noisier than the corresponding saturated plots with more scattered points of high intensity, but effective comparisons can still be drawn.

Within 0-5\angs~of the water surface (the approximate distance from the surface that \suldiox~begins interacting with the outer-most waters) the angular distribution of the single \suldiox~(in the neat-water system) is concentrated primarily in $\cos(\theta)>0$. The peak of the distribution occurs at $\cos(\theta)=1$. This indicates that the \suldiox~bisector points out of the water surface, with the sulfur atom pointing towards the aqueous bulk, and the two oxygens pointing into the gas phase. This same distribution occurs in the saturated system for positions below 5\angs. Beyond 5\angs~above the surface both distributions become isotropic. Promixity to the water highly orients the \suldiox~with the sulfur atom pointing in towards the water bulk. Moving further away from the water surface, and interacting less with \wat~molcules allows for greater orientational freedom as exhibited in the isotropy of the distributions above 5\angs. %Because both the neat-water and saturated systems show rather similar distributions, it is possible that the concentration of \suldiox~does not strongly affect the orientation of the \suldiox, unlike the orientations of waters.

@@

The plots indicate that \suldiox~orients similarly at both the low and high concentration water surface environments. The plots of $\cos(\phi)$ for both concentrations show a isotropic behavior. However, the $\theta$ values follow similar trends at both concentrations indicating that the \suldiox~sulfur orients down towards the water phase, with both oxygens pointing away from the surface once the \suldiox~is within approx. 5\angs~of the surface. This behavior continues down to at least 5\angs~below the surface, notwithstanding any chemical reactions that may occur that are not simulated using the classical MD techniques utilized in this work.

### 2011-06-07 TransitOrientation.tex
reworked the results section. One overall results section with 3 different subsections.
Finished first pass of fixing up draft 4.

\section{Steered MD Transit Simulations}

\subsection {\suldiox~Orientation}

@@

	From its starting position 20\angs~above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$. Near the surface the bisector angle $\theta$ becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase. At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems. Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface. In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface. The main difference between the two systems is where the point of the transition from isotropic to preferred orientations is found.

  % talk about theta differences
  The $\cos(\theta)$ distribution above 10\angs~from the surface is isotropic. There is a transition below 10\angs~where the \suldiox~orients with its bisector perpendicular to the water surface. On approach to the neat-water surface, that transition occurs at approximately 5\angs~from the surface. This coincides with the approximate distance of bonds formed to the outer-most water molecules. Below 5\angs, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few\angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system.

@@

  % talk about phi differences
  With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ would be isotropic relative to the reference axis. This is the case in both systems, with one exception in the neat-water system. At the neat-water surface and just below, from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens, and a clear trend appears in the profile of $\phi$. This corresponds to a \suldiox~orienting less flat to the surface (inclined up to 60 degrees from the surface normal). The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface. Figure \ref{fig:so2-surface-angles}b depicts the orientation of the \suldiox~slightly below the water surface. This is in contrast to \suldiox~molecules above the surface, oriented more perpendicularly. The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface. As the \suldiox~is solvated by more highly-coordinated bulk water, the S-O bonds experience less equal interaction environments. Baer et al. noted that their force field model for the \suldiox~does not reproduce well the first hydration shell geometries,\cite{Baer2010} so conclusions regarding the specific interactions and hydrate geometries between the \suldiox~and \wat~cannot be made here. It is notable that the same reorientation does not occur as strongly in the saturated system. The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions at the surface. Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.

>>>>

\subsection{Steered MD Transit Simulations}

\subsubsection {\suldiox~Orientation}

@@

	From its starting position 20\angs~above the water surface, until the \suldiox~moves to within 10\angs~of the surface of both systems, the orientation is isotropic in $\theta$ and $\phi$. This can be seen in the plots as mostly uniform coloration at a given distance from the surface independent of $\cos$ of $\phi$. Near the surface the bisector angle $\theta$ becomes more perpendicular ($\cos(\theta)\approx1$) with the \suldiox~sulfur pointing into the water phase. At the point when the \suldiox~reaches the surface location (distance $=0$\angs), the bisector is perpendicular in both the neat-water and saturated systems. Figure \ref{fig:so2-surface-angles}a depicts \suldiox~near the water surface with the bisector aligned perpendicular to the interface. In the absence of simulated ionic species that form through \suldiox-\wat~chemistry at the surface, it is clear that the adsorbing \suldiox~in the gas phase takes on a preferred orientation to adsorb on a water surface. The main difference between the two systems is where the point of the transition from isotropic to preferred orientations is found.

  % talk about theta differences
  Looking more closely at the approach, the $\cos(\theta)$ distribution above 10\angs~from the surface is isotropic. For the neat-water surface, a transition occurs at approximately 5\angs~from the surface where the \suldiox~orients with its bisector perpendicular to the water surface. This coincides with the approximate distance of bonds formed to the outer-most water molecules. Below 5\angs, the orientation persists and the \suldiox~begins to interact with the topmost surface waters, and orients accordingly. In the saturated system, the same trend occurs, however the onset of the perpendicular orientation begins at approximately 8\angs~above the surface. The layer of adsorbed \suldiox~already present in the saturated system most likely interacts with the transiting \suldiox~molecule. Also, topmost water molecules from the surface move up to a few \angs~inside the adsorbed \suldiox~layer and interact with the transiting \suldiox~further from the surface than those in the neat-water system. It is remarkable that the orientational trend appears so strong in the $\cos(\theta)$ plots even with so little data as was collected from the single \suldiox~molecule of each simulation. From onset of orientation above the surface until 10\angs~below, the \suldiox~holds a preferred orientation.

@@

  % talk about phi differences
	%Indeed, the $\phi$ orientations appear mostly isotropic without any clear trends for preferred orientation in that angle. This appears as a mostly uniform color across the entire plot, with minor variations that are disjointed throughout the distance range.

 With a mostly perpendicular bisector angle, it is expected that the values of $\phi$ for \suldiox~would be isotropic relative to the reference axis. This is the case in both systems, with only a few exceptions. At the neat-water surface and just below, from 0\angs~to 5\angs~in the water phase, the $\theta$ profile broadens down to $\cos(\theta)=0.5$, and a clear trend appears in the profile of $\phi$ near $\cos(\phi)=0$. This corresponds to \suldiox~inclined up to 60 degrees from the surface normal. The result is that \suldiox~in this region shows a preference for one of the S-O bonds to point more towards the aqueous bulk and the other to point away towards the interface. Figure \ref{fig:so2-surface-angles}b depicts the orientation of the \suldiox~slightly below the water surface. This is in contrast to \suldiox~molecules above the surface, oriented more perpendicularly. The behavior is likely due to the interaction between the waters and one of the S-O bonds leading to a higher solvation than above the water surface. As the \suldiox~is solvated by more highly-coordinated bulk water, the S-O bonds experience less equal interaction environments. Baer et al. noted that their force field model for the \suldiox~does not reproduce well the first hydration shell geometries,\cite{Baer2010} so conclusions regarding the specific interactions and hydrate geometries between the \suldiox~and \wat~cannot be made here. It is notable that the same reorientation does not occur as strongly in the saturated system. The presence of the adsorbed \suldiox~layer apparently decreases the reorienting behavior likely because of the disrupting effect the higher \suldiox~concentration has on the water interactions at the surface. %Below this 5\angs~region under the surface, the \suldiox~resumes an isotropic orientation as it becomes fully solvated by bulk waters.
